---
title: "Assignment 3"
author: "McKenzie Kozma"
date: "9/18/2019"
output: 
#rmarkdown::github_document
  pdf_document: default
  #html_document: default
---
```{r}
library(tidyverse)
data <- read.csv("titanic.csv")
```


REDO OF ASSIGNMENT 1 #13 - 23 (i used dplyr in A1 so the following will be done in base r)

13. Calculate the mean age of female passengers

```{r}
mean(data$Age[data$Sex == "female"], na.rm = TRUE)
```

14. Calculate the median fare of the passengers in Class 1

```{r}
median(data$Fare[data$Pclass == 1], na.rm = TRUE)
```

15. Calculate the median fare of the female passengers that are not in Class 1

```{r}
median(data$Fare[data$Sex == 'female' & data$Pclass != 1], na.rm = TRUE)
```

16. Calculate the median age of survived passengers who are female and Class 1 or Class 2

```{r}
median(data$Age[data$Survived == 1 & data$Sex == 'female' & data$Pclass %in% c(1,2)], na.rm = TRUE)
```

17. Calculate the mean fare of female teenagers survived passengers

```{r}
mean(data$Fare[data$Sex == 'female' & data$Survived == 1 & data$Age < 20 & data$Age > 12], na.rm = TRUE)
```

18. Calculate the mean fare of female teenagers survived passengers for each class

```{r}
by(data$Fare[data$Sex == 'female' & data$Survived == 1 & data$Age > 12 & data$Age < 20], 
   data$Pclass[data$Sex == 'female' & data$Survived == 1 & data$Age > 12 & data$Age < 20], FUN = mean)
```

19. Calculate the ratio of Survived and not Survived for passengers who are who pays more than the average fare

```{r}
nobs <- nrow(data[data$Fare > mean(data$Fare),])
survive <- sum(data$Survived[data$Fare > mean(data$Fare)])

survive/(nobs - survive)
```

20. Add column that standardizes the fare (subtract the mean and divide by standard deviation) and name it sfare

```{r}
sfare = (data$Fare - mean(data$Fare)) / sd(data$Fare)
data2 <- cbind(data, sfare)
```

21. Add categorical variable named cfare that takes value cheap for passengers paying less the average fare and takes value expensive for passengers paying more than the average fare.

```{r}
cfare = ifelse(data2$Fare < mean(data2$Fare), "cheap", "expensive")
data3 <- cbind(data, cfare)
```

22. Add categorical variable named cage that takes value 0 for age 0-10, 1 for age 10-20, 2 for age 20-30, and so on

```{r}
cage = trunc(data3$Age / 10) * 10
data4 <- cbind(data3, cage)
```

23. Show the frequency of Ports of Embarkation. It appears that there are two missing values in the Embarked variable. Assign the most frequent port to the missing ports. Hint: Use the levels function to modify the categories of categorical variables.

```{r}
summary(data$Embarked)
levels(data$Embarked) <- c("S", "C", "Q", "S")
```




REDO OF ASSIGNMENT 2 #4, 5, 11, 12, AND 13 

```{r}
library(readxl)
c2015 <- read_xlsx("c2015.xlsx")

dim(c2015)
set.seed(2019)
samp <- sample_n(c2015, 1000)
glimpse(samp)

samp$YEAR <- NULL
samp$TRAV_SP <- as.numeric(str_remove(samp$TRAV_SP, "MPH"))

samp %>% 
  group_by(INJ_SEV == "No Apparent Injury (O)") %>%
  summarise(speed = mean(TRAV_SP, na.rm = TRUE))

samp2 <- samp %>% 
  filter(SEAT_POS != "Front Seat, Left Side")
  
samp2 %>% 
  group_by(SEX) %>% 
  summarise(speed = mean(TRAV_SP, na.rm = TRUE))

samp2 %>% 
  group_by(DRINKING) %>% 
  summarise(speed = mean(TRAV_SP, na.rm = TRUE))
```

3. Calculate the travel speed (TRAV_SP variable) by day. Compare the travel speed of the first 5 days and
the last 5 days of months.

```{r}

samp2 %>% filter(DAY %in% c(1:5, 26:30)) %>% 
  group_by(DAY %in% c(1:5)) %>% 
  summarise(speed = mean(TRAV_SP, na.rm = TRUE))

#drive faster at the end of the month
```

4. Calculate the travel speed (TRAV_SP variable) by day of the week. Compare the travel speed of the
weekdays and weekends.

```{r}
samp2 %>% group_by(DAY_WEEK %in% c("Sunday", "Saturday")) %>% 
  summarise(speed = mean(TRAV_SP, na.rm = TRUE))

#faster on weekends
```

5. Find the top 5 states with greatest travel speed.

```{r}
samp2 %>% group_by(STATE) %>% 
  summarise(SPEED = mean(TRAV_SP, na.rm = TRUE)) %>% 
  top_n(5, SPEED)
```

6. Rank the travel speed by MONTH.

```{r}
samp2 %>% group_by(MONTH) %>% 
  summarise(SPEED = mean(TRAV_SP, na.rm = TRUE)) %>% 
  arrange(-SPEED)
```

7. Find the average speed of teenagers in December.

```{r}
samp2 %>% filter(MONTH == "December" & AGE %in% c(13:19)) %>% 
  summarise(SPEED = mean(TRAV_SP, na.rm = TRUE))
```

8. Find the month that female drivers drive fastest on average.

```{r}
samp2 %>% filter(SEX == "Female") %>% 
  group_by(MONTH) %>% 
  summarise(SPEED = mean(TRAV_SP, na.rm = TRUE)) %>% 
  top_n(1, SPEED)
```

9. Find the month that male driver drive slowest on average.

```{r}
samp2 %>% filter(SEX == "Male") %>% 
  group_by(MONTH) %>% 
  summarise(SPEED = mean(TRAV_SP, na.rm = TRUE)) %>% 
  top_n(-1, SPEED)
```

10. Create a new column containing information about the season of the accidents. Compare the percentage
of Fatal Injury by seasons.

```{r}
samp2 %>% 
  mutate(SEASON = ifelse(MONTH %in% c("December", "January", "February"), "Winter",
                         ifelse(MONTH %in% c("March", "April", "May"), "Spring",
                                ifelse(MONTH %in% c("June", "July", "August"), "Summer",
                                       ifelse(MONTH %in% c("September", "October", "November"), "Fall", ""))))) %>% 
  group_by(SEASON) %>% 
  summarise(prop_fatal = (sum(INJ_SEV == "Fatal Injury (K)"))/n())
  
```

11. Compare the percentage of fatal injuries for different type of deformations (DEFORMED variable)

```{r}
samp2 %>% group_by(DEFORMED) %>% 
  summarise(prop_fatal = (sum(INJ_SEV == "Fatal Injury (K)"))/n()) %>% 
  arrange(-prop_fatal)
```

